Experience Report:

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

\section{API}

The main idea is to start with a generic function, with the
following interface: 
% TODO: rename this function
\begin{ocaml}
  val compute : 
    f:('a -> 'b) -> 
    handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
list of \emph{tasks} of type \ocaml{('a * 'c) list}.
The function \ocaml{f} is applied to the first component for each task.
Whenever it completes, the task and the result are passed to the
function \ocaml{handle}. This may possibly generate a new list of
tasks. The process ends when all computations are done.

@@

\section{Derived API}

@@

\section{Implementation Details}

cores?

network: protocol, fault tolerance, etc.

\section{Experiments}

@@

\section{Future Work}

>>>>

\section{API}\label{sec:API}

The main idea is to start with a generic function, with the
following interface: 
% TODO: rename this function?
\begin{ocaml}
  val compute : 
    f:('a -> 'b) -> 
    handle:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
% TODO be even more explicit about the introduction of the word `task'
This function takes two functions, \ocaml{f} and \ocaml{handle}, and a
list of \emph{tasks} of type \ocaml{('a * 'c) list}.
The function \ocaml{f} is applied to the first component for each task.
Whenever it completes, the task and the result are passed to the
function \ocaml{handle}. This may possibly generate a new list of
tasks. The process ends when all computations are done.

@@

\section{Derived API}\label{sec:derived}

@@

These five functions are the most natural derivations of the
\of{compute} functions. There are obviously other possible derivations
(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
in most cases, they could be derived in a straightforward way from our
generic API.

% TODO cannot be instantiated in the monomorphic case

\section{Implementation Details}\label{sec:implem}

We now describe the implementations of the various modules introduced
in Section~\ref{sec:API}. The implementation of the \of{Sequential}
module is straightforward and does not require any explanation.

\subsection{Multiple Cores}

The \of{Cores} module implements the distributed computing library for
multiple cores on the same machine. It provides a function
\of{set_number_of_cores: int -> unit} to indicate the number of cores
to be used. The number passed to this function may actually be
different from the actual number of cores in the machine; it is rather
the number of tasks to be performed simultaneously. 
% TODO comment more on the number of cores?
Thus a program using the \of{Cores} module typically begins with 
\begin{ocaml}
open Mapreduce.Cores
let () = set_number_of_cores 8
\end{ocaml}
% TODO rename Mapreduce?

The \of{Cores} module is implemented with \unix\ processes, using the
\of{fork} and \of{wait} system calls provided by the \of{Unix} library
of \Ocaml. The idea is pretty simple. 
The \of{compute} function maintains a global table of pending tasks
and keeps track of the number of idle cores.
Whenever there a pending task and an idle core, 
a new sub-process is created using \of{Unix.fork}; 
once completed, the sub-process marshals the result 
into a local file. The main process maintains a table
mapping each sub-process ID to the input task and the local file
name. It waits for any completed sub-process using \of{Unix.wait}
and recovers the result from the local file. Then function \of{handle}
is applied, which may generate new tasks.
The main loop can be depicted in the following way:
\begin{flushleft}
  \quad  \textbf{while} pending tasks $\lor$ pending sub-processes \\
  \quad  \quad \textbf{while} pending tasks $\land$ idle cores \\
  \quad  \quad \quad create new sub-process for some task \\
  \quad  \quad \textbf{wait} for any completed sub-process \\
  \quad  \quad \quad push new tasks generated by handle \\
\end{flushleft}

It would be more efficient to use \unix\ pipes instead of local files
but pipes are severely restricted w.r.t. data size. To be more
generic, we make no assumption about the size of the data. Providing
an alternative implementation using pipes would be straightforward.

The scheduling of tasks to the different cores is left to the
operating system, through \of{Unix.fork}. Thus it may be the case that
two tasks are executed on the same core, even if the declared number
of cores is less or equal than the actual number of cores on the machine.
% TODO: comment on user-provided scheduling?

\subsection{Network of Machines}

The \of{Network} module implements the distributed computing library for
a network of machines. 
It provides a function \of{declare_workers: n:int -> string -> unit} 
to fill a table of worker machines. The argument \of{n} indicates the
number of cores to be used in each worker machine, analogous to 
\of{Cores} module. Thus a program using the \of{Network} module
typically begins with
\begin{ocaml}
open Mapreduce.Network
let () = declare_workers ~n:12 "machine1"
let () = declare_workers ~n:8  "machine2"
...
\end{ocaml}

The \of{Network} module is based on a traditional client/server
architecture, where each worker is a server and the master is the
client of each worker. The main execution loop is similar to the one
in the \of{Cores} module, where distant processes on remote machines
correspond to sub-processes and idle workers correspond to idle
cores. In addition to this, it also involves issues of message
transfer and fault tolerance, which are subsequently described.

\subsubsection{Protocol}

\subsubsection{Fault Tolerance}\label{sec:fault}

\section{Experiments}\label{sec:experiments}

@@

\section{Future Work}\label{sec:future}
